---
title: "BST 260 Problem set 3: Logic, reading, guns and geography"
author: "Yixuan Xin"
format: html
editor: visual
---

### Exercise 1.

- $C$ = the coupon is valid

- $P$ = the price is reduced

- $S$ = the shipping is free

The statement can be formalized as

$$C \rightarrow (P \lor S)$$

It is false when $C$ is true and both $P$ and $S$ are false.

### Exercise 2.

- S = it is sunny

- W = it is warm

- P = picnic is cancelled

$$ (\neg S \wedge \neg W)\rightarrow P$$

It is false if it is not sunny and not warm, but the picnic is not cancelled.

### Exercise 3.

- E = you may enter the contest

- M = you are a member

- P = you have paid the fee

$$E\leftrightarrow (M \lor P)$$

If you enter but the sentence is false, then we should know that you are not a member and you did not pay the fee. If E is true, then $M \lor P$ should be false.

### Exercise 4.

This issue is that this data is from 2010, which is too old to demonstrate current patterns. Moreover, it is not a primary dataset source because it is from Wikipedia instead of an original FBI source.

### Exercise 5.

```{r}
dim(cdcmurd)
```

> \[1\] 612 5

The output shows that there are 612 rows in the dataset. Adjusted rate definition makes sense because it allows us to compare death rates across different age distributions.

### Exercise 6.

One will use the command table(table(head(cdcmurd\$Location,120))) if he wants to know the number of times the states' location appear. The output gives me:

2 3

33 18

It shows that among the first 120 rows, there are 33 states' locations appear twice and 18 states' locations appear three times.

The results from console give:

```{r}
table(table(head(cdcmurd$Location,153)))
```

```{r}
head(table(cdcmurd$Location))
```

> AK AL AR AZ CA CO
>
> 12 12 12 12 12 12

For the first one, it means there are 51 states' locatins appear three times in the first 153 rows, and the second one is the number of times each states' names appear. To look at the second one clearer, I use:

```{r}
table(table(cdcmurd $Location))
```

> 12 51

The result shows that for 51 states each location appears 12 times, which is corresponding to 51(12)=612 rows.

From this, I learn that the table(table()) construction is good to show the frequency of number of appearance of each category.

### Exercise 7.

any(is.na(x) tries to find any missing value and from a series of boolean and return true if there is at least one missing value and vise versa. function(x) is a function. apply() function wants to find any missing value in cdcmurd for each column(margin=2). In general, this line of code wants to check each column of cdcmurd and see if there is at least one missing value in that column.

```{r}
apply(cdcmurd,2,function(x) any(is.na(x)))
```

YEAR Location Death.Rate Deaths URL \> \> FALSE FALSE TRUE FALSE FALSE

The output shows that for Death.Rate column, it has at least one missing value.

### Exercise 8.

This line gets all rows that contains an NA in the Death.Rate column in the cdcmurd, and count how many rows are there in total.

```{r}
cdcmurd[which(is.na(cdcmurd $Death.Rate)),] |> nrow()
```

> \[1\] 28

The result shows there is 28 rows having missing Death.Rate values.

### Exercise 9.

```{r}
cdcmurd[which(is.na(cdcmurd $Death.Rate)),] |> nrow()
```

> \[1\] 28

They both have same result of 28, representing 28 rows having missing values in Death.rate column. Code in exercise 8 gives a numeric value, while code in exercise 9 gives a dataframe and one column called n contains 28.

```{r}
median(cdcmurd$Deaths)
```

> \[1\] 250

### 

```{r}
cdcmurd |> filter(is.na(Death.Rate)) |>  group_by(Location) |> summarise(mn=mean(Deaths))
```

> \# A tibble: 6 × 2 Location mn
>
> <chr> <dbl>
>
> 1 ME 19\
> 2 ND 14.5
>
> 3 NH 17.5
>
> 4 RI 19\
> 5 VT 13.7
>
> 6 WY 17.2

From the result, we see that the mean value among these states having missing values have a relatively low mean deaths, ranging from 10 to 19. However, the median number of deaths is 250, which is higher. Therefore, the death rates are missing probably because a low death numbers and cdc may not report a statistically stable death rate. I learn to compare different summary statistics and compare there differences.

### Exercise 10.

- Line 1 will filter 204 rows.

- For Line 2, Death.Rate define the y coordinates of the plot.

- the boxplot shows the distribution of a continuous variable and some statistics, including the median, two hinges, two whiskers, and outlying points.The geom_beeswarm gives offset points in categories to avoid overplotting.

  - Output from R documentation:

    Boxplot Description:

    The boxplot compactly displays the distribution of a continuous variable. It visualises five summary statistics (the median, two hinges and two whiskers), and all "outlying" points individually.

  - geom_beeswarm:

    The beeswarm geom is a convenient means to offset points within categories to reduce overplotting. Uses the beeswarm package.

  - Output from chatbot: The boxplot provides a statistical summary of the distribution of `Death.Rate` for each year. It shows the median, the first and third quartiles, the interquartile range, whiskers, and potential outliers. The beeswarm plot, in contrast, shows the individual observations. Each point represents a location’s death rate, and points with similar values are shifted horizontally to reduce overlap. Thus, the boxplot emphasizes summary statistics, while the beeswarm preserves information about the individual observations and the shape and density of the distribution.

  ```{r}
  myp2 <- cdcmurd |>
    filter(YEAR %in% c(2019, 2020, 2021, 2022)) |>
    ggplot(aes(factor(YEAR), Death.Rate, text = Location)) +
    geom_boxplot(outlier.shape = NA) +
    geom_beeswarm(alpha = 0.7) +
    labs(
      x = "Year",
      y = "Homicide death rate per 100,000",
      title = "Homicide Death Rates by State, 2019–2022"
    )
  ggplotly(myp2)
  ```

  The ubdated version contains a better X-axis name called Year, and the Death.Rate specifies the unit of death, and it adds a descriptive title. Moreover, it makes the points less dense and visually straightforward.

### Exercise 11.

- 11.a: download.file(url, destfile). It can be more interpretable through separating the creation of the temporary file and the download operation to two lines.

  tf \<- tempfile() download.file("[https://tinyurl.com/randgunlaw",](https://tinyurl.com/randgunlaw%22,) tf)

- 11.b : read_xlsx() is for reading excel files,with a path as input. 2 means get the second worksheet.

- 11.c: there are 1902 rows.

- 11.d: 1813.

- 11.e: I use the chatgpt model and my prompt is: use dplyr and retrieve the earliest value fo the Effective Date Year column from the lawdat data frame. I learn that summarize is a brief way to sort and select.

  - Code and result:

  ```{r}
  lawdat |>
    summarize(
      earliest_year = min(`Effective Date Year`, na.rm = TRUE)
    )
  ```

  > 1813

- 11.f.1: The take home message is the most common change is the implementation of the restrictive gun laws(1241), and the least common change is 'see note' with a permissive effect(2).

- 11.f.2:

```{r}
lawdat |>
  count(`Type of Change`, Effect, sort = TRUE)
```

> 1 Implement Restrictive 1241
>
> 2 Modify Restrictive 215
>
> 3 Modify Permissive 192
>
> 4 Implement Permissive 187
>
> 5 Repeal Permissive 59
>
> 6 Repeal Restrictive 6
>
> 7 See note Permissive 2

From the code, we can see that this result matches with the reference code. I think count() is a brief way to group variables and sort is helpful for listing from the most to the least.

### Exercise 12.

The code first filter the column 'Effect' with only the 'Permissive', and then use summarize to group by 'Effective Date Year', and calculate how many permissive legislative changes happen in that year. Then it uses ggplot() to put the x axis as 'Effective Date Year'. and y axis as the number of changes, and making a dot plot.

It cannot associate each point with certain state because we only group by Effective Date Year instead of by state. Each point represents legislative changes for multiple states instead of only one.

### Exercise 13.

Missouri and Tennessee have the largest frequencies of legislative changes that have “Permissive” effects(15). California state have the largest frequencies of legislative changes that have “Restrictive” effects(73).

```{r}
lawdat |>
  count(Effect, State) |>
  slice_max(n, by = Effect)
```

> \# A tibble: 3 × 3 Effect State n <chr> <chr> <int>
>
> 1 Permissive Missouri 15
>
> 2 Permissive Tennessee 15
>
> 3 Restrictive California 73

### Exercise 14.

The choice of states: As the exercise 13 shows, California has the largest frequencies of legislative changes that have 'Restrictive' effects, and Tennessee has the largest frequencies of legislative changes that have “Permissive” effects. New York has more restrictive changes over permissive changes with a big difference. These states allow for comparison of homicide-rate patterns among states with different changes. Mississippi is chose probably because it has a more balanced permissive and restrictive.

Patterns: From the plot we can see that New York has the lowest death rate, and then California, Mississippi, and Tennessee is the highest. The four states all show general similar pattern across years. Mississippi has consistenly higher highest death rate than other states, and the highest is around 2021.

Data filtering and combination: we can harmonize the state identifiers and combine RAND legislation data with the CDC homicide data by state and year. Then we can compare the type of legislative changes with death rates and define an appropriate time period.
